---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Abstract

Below paper was prepared for the needs of the final project for Advanced Econometrics at the University of Warsaw. By using ``Multinomial Logit Model`` on the data of Titanic survivors
we wanted to verify whether the socio-economic status or paying higher fare impacted passenger's probability of survival. The data is originally available on the ``Kaggle`` competition website "Titanic: Machine Learning from Disaster" and contains  data for 1.309 passengers indicating whether they survive, what was their material status, what gender they were, what age they were etc. As our explanatory variables are individual specific (they do not change across alternatives) we decided to use ``Multinomial Logit Model``. In this paper, dependent variable ``Survival`` was explained with different independent variables that were selected using multinomial logit model and verified by statistical tests. The econometric model was built in ``R`` with the use of ``??? packages``. The end result is a final model with significant variables that best explains survival rate.

__Keywords:__ Titanic, Survival Rate, Multinomial Logit Model.

# 1. Introduction

The sinking of the Titanic is one of the most infamous shipwrecks in history. On April 15, 1912, during her maiden voyage, the widely considered “unsinkable” RMS Titanic sank after colliding with an iceberg. Unfortunately, there were not enough lifeboats for everyone onboard,resulting in the death of 1502 out of 2224 passengers and crew. While there was some element of luck involved in surviving, it seems some groups of people were more likely to survive than others. Therefore this topic seems interesting for further investigation to see which variables influenced survival rate the most.

We will verify the following hypothesis using ``multinomial logit model``:


__Hypothesis 1:__


H0: Having a seat in higher class significantly increases the chance of passenger to survive.  

H1: Having a seat in higher class did not increases the chance of passenger to survive.

__Hypothesis 2:__

H0: Paying a higher fare/having a family member on board increases the chance of passenger to survive.
  
H1: Paying a higher fare/having a family member on board does not increase the chance of passenger survival.


# 2. Literature review


# 3.Dataset and preprocessing

## 3.1.Dataset

## 3.2.Preprocessing

```{r}
library("sandwich")
library("zoo")
library("lmtest")
library("MASS")
library("aod")
library("nnet")
library("Formula")
library("miscTools")
library("maxLik")
library("mlogit")
library("car")
library("survival")
library("AER")
library("stargazer")
```




```{r}
dfx <- read.csv("train.csv")
df1 <- read.csv("test.csv")
df2 <- read.csv("gender_submission.csv")
dfy <- merge(df1, df2)
df <- rbind(dfx, dfy)
```


```{r}
head(df)
```


```{r}
df <- subset(df, df$Embarked != "")
df <- subset(df, df$Fare != "")
df$Survived <- ifelse(df$Survived == 0, "No", "Yes")
df$Pclass <- as.factor(df$Pclass)
df$Embarked <- as.factor(df$Embarked)
df$Survived <- as.factor(df$Survived)
df$Sex <- as.factor(df$Sex)
head(df)
```


```{r}
summary(df)
```



```{r}
df$Family <- df$SibSp + df$Parch
head(df)
```



```{r}
fdf <- df
fdf <- fdf[,c("Sex", "Pclass", "SibSp", "Parch", "Family", "Embarked", "Survived", "Fare")]
fdf$Farepp <- fdf$Fare/(fdf$Family + 1)
```

```{r}
model1 <- multinom(Survived ~ ., data = fdf)
model2 <- multinom(Survived ~ .-Fare - Farepp, data = fdf)
model3 <- multinom(Survived ~ .-Fare - Farepp-Embarked, fdf)
model4 <- multinom(Survived ~ .-Fare - Farepp-Embarked-SibSp-Parch, fdf)
```


```{r}
summary(model1)
summary(model2)
summary(model3)
summary(model4)
```

log(P(Survived=Yes)/P(Survived=No)) = 3.07093350 + (-3.73011354 * Sexmale) + (-0.75640797 * Pclass2) + (-1.64138819 * Pclass3) + (-0.13429952 * SibSp) + (0.04530608 * Parch) + (-0.08899344 * Family)
+ (-0.03205837 * EmbarkedQ) + (-0.31817849 *EmbarkedS)


```{r}
# statistical significance
z1 <- summary(model1)$coefficients/summary(model1)$standard.errors
z2 <- summary(model2)$coefficients/summary(model2)$standard.errors
z3 <- summary(model3)$coefficients/summary(model3)$standard.errors
z4 <- summary(model4)$coefficients/summary(model4)$standard.errors
```



```{r}
# 2-tailed z test
p1 <- (1 - pnorm(abs(z1), 0, 1)) * 2
p2 <- (1 - pnorm(abs(z2), 0, 1)) * 2
p3 <- (1 - pnorm(abs(z3), 0, 1)) * 2
p4 <- (1 - pnorm(abs(z4), 0, 1)) * 2
```

```{r}
stargazer(z1,z2,z3,z4, type = "text")
```

```{r}
stargazer(p1,p2,p3,p4, type = "text")
```

```{r}
stargazer(model1, model2, model3, model4, type = "text")
```



```{r}
fdf <- df[,c("Survived", "PassengerId", "Pclass", "Sex", "SibSp", "Parch", "Embarked", "Family", "Fare")]
fdf$Farepp <- df$Fare/(df$Family+1)
mldf <- mlogit.data(fdf, shape = "wide", choice= "Survived", v.names=c("PassengerId"))
```

```{r}
model <- mlogit(Survived~ 0| Sex + Family + Pclass, data = mldf)
model
```

```{r}
summary(model)
```
```{r}

#model <- mlogit(Survived~ Fare+Farepp | Sex + Family + Pclass, data = mldf)
#model

```





```{r}
# a) general-to-specific method for variables selection




```



```{r}
# b) at least one nonlinear relationship (variable to a power) and interaction between variables

fdf$Fare2 <- fdf$Fare^2
fdf$Farepp2 <- fdf$Farepp^2
nlr_model1 <- multinom(Survived ~ Fare + Farepp + Fare2 + Farepp2, data = fdf)
nlr_model2 <- multinom(Pclass ~ Fare + Farepp + Fare2 + Farepp2, data = fdf)

summary(nlr_model1)
summary(nlr_model2)


# ? Interactions 

```


```{r}
# c) calculation and interpretation of marginal effects for the final model (from the general-tospecific approach)



model<-mlogit(Survived~0|Fare+Family+Sex, data=mldf)
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}

mldf$Sex

#mldf$Sex[mldf$Sex=="male"] <- 1
#mldf$scode[mldf$Sex=="female"] <- 0

mldf$Sex

z <- with(mldf, data.frame(Sex=tapply(Sex, index(model)$alt, mean),
                            disp=tapply(Fare, index(model)$alt, mean),
                            price=tapply(Family, index(model)$alt, mean)))
z[,1:2] = 1
z



```



```{r}
#present the general model, the final model (the specific model) in one quality table. If there is space, at least one intermediate model might be presented


# Easy in the end using stargazer

```


```{r}
linktest = function(model) {
  # written by dr Rafal Wozniak, Faculty of Economic Sciences, University of Warsaw
  # 2019-04-18
  #
  # arguments:
  # ------------------
  # model - model estimated by glm function
  
  # check if it is of class 'glm'
  
  # Linktest
  y = model$y
  yhat = log(model$fitted.values/(1-model$fitted.values))
  yhat2 = yhat^2
  # auxiliary regression
  aux.reg = glm(y~yhat+yhat2, family=binomial(link=model$family$link))
  show(summary(aux.reg))
  return(aux.reg)
}
```


```{r}
# e) perform the linktest and interpret the result
mylogit <- glm(Survived~ Sex + Pclass, data=fdf, family=binomial(link="logit"))
linktest(mylogit)
```


